# K-means
`r Sys.Date()` updated  

K-means clustering: It is a method to cluster n points into k clusters based on the means of shortest distance. (cautsion: do not confused it with K-nearest clustering). We use the demo data sets "USArrests".

```{r, warning=FALSE}
data("USArrests")
df <- USArrests
head(df)
```
```{r}
# scaling the data
df <- scale(df)
head(df)
```
We use **factoextra** package to create beautiful clusters visualization

We also use `fviz_nbclust` to determine the optimal number of clusters
```{r}
set.seed(123)
fviz_nbclust(df, kmeans, method="wss") + geom_vline(xintercept = 4, linetype = 2)
```

```{r}
set.seed(123)
(km.res <- kmeans(df, centers = 4, nstart = 25))
```

There are 7 Avaliable components you can access it.    
Compute the mean of each variables by clustering using the original data:
```{r}
aggregate(USArrests, by=list(cluster=km.res$cluster), mean)
```
Display results with your cluster result:
```{r}
dd <- cbind(USArrests, cluster = km.res$cluster)
head(dd)
```
Plot the clustering result using factoextra package
```{r, message=FALSE}
library(factoextra)
fviz_cluster(km.res, df, palette="Set2", ggtheme = theme_minimal())
#points(km.res$centers, col = 1:2, pch = 8, cex=2)
```

Reference: [here](https://www.datanovia.com/en/lessons/k-means-clustering-in-r-algorith-and-practical-examples/)

