# K-means
`r Sys.Date()` updated  

## Case 1
K-means clustering: It is a method to cluster n points into k clusters based on the means of shortest distance. (cautsion: do not confused it with K-nearest clustering). We use the demo data sets "USArrests".

```{r, warning=FALSE}
data("USArrests")
df <- USArrests
head(df)
```
```{r}
# scaling the data
df <- scale(df)
head(df)
```
We use **factoextra** package to create beautiful clusters visualization

We also use `fviz_nbclust` to determine the optimal number of clusters
```{r}
set.seed(123)
fviz_nbclust(df, kmeans, method="wss") + geom_vline(xintercept = 4, linetype = 2)
```

```{r}
set.seed(123)
(km.res <- kmeans(df, centers = 4, nstart = 25))
```

There are 7 Avaliable components you can access it.    
Compute the mean of each variables by clustering using the original data:
```{r}
aggregate(USArrests, by=list(cluster=km.res$cluster), mean)
```
Display results with your cluster result:
```{r}
dd <- cbind(USArrests, cluster = km.res$cluster)
head(dd)
```
Plot the clustering result using factoextra package
```{r, message=FALSE}
library(factoextra)
fviz_cluster(km.res, df, palette="Set2", ggtheme = theme_minimal())
#points(km.res$centers, col = 1:2, pch = 8, cex=2)
```

Reference: [here](https://www.datanovia.com/en/lessons/k-means-clustering-in-r-algorith-and-practical-examples/)  

## Case 2
Use iris data for a second demo
```{r}
str(iris)
```
We use the features with "Length" to do a kmeans clustering:
```{r}
i <- grep("Length", names(iris))
x <- iris[, i]
cl <- kmeans(x, centers = 3, nstart = 10)
plot(x, col = cl$cluster)
```

## Hierarchical clustering
```{r}
d <- dist(iris[, 1:4]) # calculate the euclidean distance
hcl <- hclust(d)
hcl
```
```{r}
plot(hcl)
```

Usually, you will need to cut the tree to define the clusters with `cutree()`.
```{r}
(cutree(hcl, h=1.5)) # at certain height
cutree(hcl, k=2) # with certain number of clusters
```
```{r}
plot(hcl)
abline(h = 3.9, col="blue")
```

You can check the result using `identical()`
```{r}
identical(cutree(hcl, k=3), cutree(hcl, h=3.9))
```

